Low prevalence of \emph{a priori} power analyses in motor behavior research

### 2022-10-19 docs/revision-1/r1_response-to-reviewers.Rmd

---
title       : "Low prevalence of *a priori* power analyses in motor behavior research"
#authors     : "Brad McKay, Abbey Corson, Mary-Anne Vinh, Gianna Jeyarajan, Chitrini Tandon, Hugh Brooks, Julie Hubley, & Michael J. Carter"
journal     : "Journal of Motor Learning and Development"
manuscript  : "JMLD.2022-0042"

class       : "final"
output      : papaja::revision_letter_pdf
---

Dear Dr. Zelaznik,

Thank you overseeing the review of our manuscript for publication at the _`r rmarkdown::metadata$journal`_. We appreciate the positive and helpful comments we received from this review process. We believe these comments have helped to strengthen our manuscript.

Based on comments from both Reviewers, as well as yourself, we have made some changes to the manuscript. Below we provide out point-by-point response (in normal font) to the Reviewers' comments (in ***bold-italic***). All substantial changes in the revised manuscript are included in this response letter (in a text box) and appear in \textcolor{blue}{blue} text in the revised manuscript.

We would like to draw your attention to a change in the revised manuscript based on an error we recently discovered. We found that an article from 2017 was included in the sample due to a faulty DOI error. This article included a power analysis so our reported numbers in the revised manuscript have changed slightly for most analyses. We have also updated Figure 1 to reflect the removal of this paper.

Thank you again for considering our work and overseeing the review process.

# Editor comments

\RC{\emph{I have one suggestion of my own. As you examined three journals, but you do not present any ``effect of journal'' reporting. I think you should. Think about it. I won't demand this revision.}}

Thank you for this suggestion. We have added an analysis comparing the proportion of studies with a power analysis in each journal to the paper in an Exploratory analysis section. The following have been added to the manuscript:

> Page 11, Lines 173-175:
>
> Exploratory analysis
>
> We used a 3-sample proportion test to compare the proportion of studies including a power analysis in the three journals surveyed.

> Page 15, Lines 221-226:
>
> Exploratory analysis
>
> Difference among journals in proportion of studies with a power analysis
> 
> Studies published in Human Movement Science included a power analysis (59/271) 18% of the time, the Journal of Motor Learning and Development (9/128) 10% of the time, and the Journal of Motor Behavior (16/152) 7% of the time. The 3-samples proportions test was significant, $\chi^2$ = 13.52, *df* = 2, *p* = .001.

\Done

# Reviewer 1 comments

\RC{\emph{I have thoroughly read this manuscript and am familiar enough to make comment. For transparency, I should also note that I am familiar with the preprint version of this manuscript.}

\emph{Currently, the manuscript is strong and so my comments are relatively minor and focus on readability and clarity. For organizational purposes, I present my comments in order of appearance within the manuscript.}}

Thank you for the positive comments about our manuscript.

\RC{\emph{Page 3, line 44. The authors might consider adding a statement indicating that the purpose of a pilot study should not be to determine estimates of the true effect, and any estimates reported from the underpowered sample are unreliable.}}

We generally agree with this comment for the reasons contained in the manuscript. Since we are arguing that estimates from previous studies and even meta-analyses are not the ideal input for a power analysis, we decided to refrain from criticizing pilot studies more forcefully than we already do.

\RC{\emph{Page 4, line 65. Should "a power analysis" be "an a priori power analysis"?}}

Unfortunately, it was rarely specified whether the power analyses that were reported were conducted a priori. We anticipated this before beginning and chose not to restrict our coding to explicitly a priori analyses.

\RC{\emph{The authors do an excellent job throughout the manuscript with "hand holding" and explaining the implication of earlier research and issues with power.}}

Thank you.

\RC{\emph{Page 5, lines 91-93. This might be a point for the discussion, but one reason for a lack of difference here is that some people perform "a post-hoc a priori power analysis". That is, people might report a priori power analyses that fit their final sample size, rather than recruit a sufficient sample size to fulfill their true a priori power analysis. I know that is cynical. But, I hypothesize it happens sometimes, given resource constraints, incentives to publish, journal pressures, etc. Could this partly explain some of the results?}}

We do not think Reviewer 1 is being cynical by suspecting some of the power analyses were conducted after the fact. We think this happens too. We have added the paragraph below to address the myriad selection effects that may be involved in the data generating process that created our sample.

> Pages 18-19, Lines 305-311
>
> Selection bias makes it challenging to interpret the impact of power analyses on positive results and sample sizes. It is possible that power analyses are sometimes added to a
study to defend a null result or a small sample size. It is also possible that researchers who conduct power analyses choose to study larger effects, leading to smaller sample sizes. We encourage readers to focus on the overall low frequency of power analysis reporting, small average sample sizes, and high rates of positive results as a description of the modern motor behavior literature.

\RC{\emph{Table 1. "Calculate average per group". Does this mean the number per group? I apologize if I missed it, but I am not sure what average means in this context.}}

We only extracted one value, so in case of an unbalanced design we took the average number per group.

\RC{\emph{Figure 2 is excellent.}}

Thank you.

\RC{\emph{Page 16, lines 225-227. I think also including the percentages of each type of analysis in the discussion would be useful: For example, 63\% of studies that reported a power analysis based their effect size assumption on a previous study (n, \%), a pilot study (n, \%), or on effect size benchmarks (n, \%).}}

Thank you for this suggestion. We have added these values. Please see Page 16, lines 240-241.

\RC{\emph{Page 16, line 251. Maybe more explicitly explain what censored means here? For example ``... such that many small effects or null results never made it into the pages of a journal or the scientific record due to publication biases...''.}}

We have added the following footnote to clarify our usage of censored:

> Page 17
>
> We use the term "censored" here to mean that some experiments fail to enter the published literature for results-dependent reasons, commonly called publication bias or reporting bias.

\RC{\emph{I hope that the authors find these comments to be useful.}}

We appreciate Reviewer 1’s feedback, which has helped to improve our manuscript.

# Reviewer comments

\RC{\emph{I would like to declare upfront that I am aware of the authors of this manuscript. I say this since it seems the journal operates a double-blind review procedure. I have seen this work presented and had already read the preprint of the manuscript prior to conducting the review.}

\emph{I commend the authors on an important and rigorously conducted study. This topic has been gathering more and more attention in the broad movement science field and this specific focus on three motor behavior journals expands on previous analyses and draws more attention to the issues. It is also promising that the Journal of Motor Learning \& Development considers it for publication since self-acknowledgement of these issues by the entities with the power to make changes (journals) will hopefully be impactful going forward. I appreciate that the authors have shared their materials and code on the OSF. Since I do note code with R, I have not checked the code but the fact that it is publicly shared already gives me confidence in the transparency and robustness of the work. In my review, I did not detect any specific errors or issues in the numerical results. I have made a few specific comments for the authors below for their consideration but overall, I have no major concerns or comments on the manuscript. One thing that could be addressed more in the discussion is how to tackle this issue. The authors stop at the enforcement of journal guidelines, but I am not sure that will be enough for the reasons I outline below. Perhaps the authors can expand further on that section to give more direction for others in the field.}}

We would like to thank Reviewer 2 for their feedback during the review process, which has helped to improve the quality of the manuscript.

\RC{\emph{Lines 47-49: It might also be worth noting that using an effect size estimate from a meta-analysis, even when accurate, results in a power analysis for the ``average'' effect in the literature, which may or may not be what the researcher really wants to be able to reliably detect.}}

This is a great point. We have added the following text to expand on this:

> Page 4, Lines 52-55
>
> Targeting the estimated mean effect will only provide the desired power if the effect is average or above average in the specific context being investigated. Therefore, even if researchers are privy to a precise and unbiased estimate of the mean effect they are studying, they still might choose a different effect to target in their power analysis.

\RC{\emph{Line 91-93: I am curious what the thought behind this hypothesis was, especially considering that no specific direction of the effect was stated. The authors write that the field is typically underpowered, so why would the presence of a power analysis result in smaller sample sizes? Another aspect here is that this assumes that the power analyses are correctly performed, with a justifiable effect size of interest and not “gamed” or adjusted to reach a more feasible sample size. The recent article on power analyses in the journal Gait and Posture found that there was no clear change in sample sizes after the journal introduced a requirement for articles to report their power analyses, which the authors also suggested might be due to “power analysis hacking” to achieve more feasible sample sizes.}}

We did not specify a direction because of the myriad factors that impact our access to the sample. In our sample, anything from "power analysis hacking" to researchers choosing different research questions that could be addressed with smaller samples might have led to smaller sample sizes being associated with reporting a power analysis. We can imagine a context in which we would predict that power analyses lead to larger samples: if we randomly assigned researchers to conduct a motor behaviour experiment with or without conducting an appropriate power analysis, granted them unlimited time and resources, and required them to complete the experiment. Instead, due to selection bias we decided to avoid making any confirmatory predictions and instead focused on describing the literature.

\RC{\emph{Lines 121-122: Did a hypothesis have to be explicitly stated (i.e., at the end of the introduction) or did the presence of statistical hypothesis testing suffice?}}

We extracted a quote of each hypothesis, so it had to be explicitly stated. That said, the hypothesis did not need to be precisely stated. For example, if the authors wrote "We tested whether our [independent variable] had an effect on our [dependent variable]" we would consider that a hypothesis because it implies a test of the null hypothesis.

\RC{\emph{Line 130: Eight people is quite a large group for data extraction. Was agreement in screening and extraction assessed? In what ways was the consistency of this process ensured?}}

During piloting the research team extracted data from 40 articles. There were 4 disagreements out of 120 items. We describe this detail in our preregistration, but can add it to the manuscript if necessary). Beyond this initial assessment, we relied on the fact each coded value was agreed on by two researchers, whether initially or after a third researcher resolved the conflict. As a final layer of oversight, the researchers flagged articles they found challenging, prompting the intervention of the lead author. The challenging papers were used as learning opportunities and therefore received the most attention.

\RC{\emph{Line 223-224: I would disagree with this conclusion slightly. The results suggest that the reporting of power analyses is not yet widely adopted. It is not possible, with the current data, to determine whether or not researchers use power analyses in planning their studies. Indeed, many ethics review boards require this, so there may be some percentage of the articles that did not report a power analysis had actually previously performed one.}}

We did not intend for this conclusion to come across so strongly. We have changed our wording to the following:

> Page 15, Line 237
>
> Our results cause us to wonder whether motor behavior research has not yet widely adopted power analyses to inform study design.

\RC{\emph{Lines 252-257: This finding is similar to the Gait and Posture paper – the guidelines explicitly require sample size justification, but the majority of articles don’t report one. Stricter enforcement and perhaps reviewer guidelines would perhaps be beneficial. The broader problem here though is that this will also encourage post-hoc power analyses if authors have not actually performed an a priori power analyses. So while such enforcement might increase the percentage of papers with a power analyses, at least a portion of these will have been conducted after the study was conducted. A stricter requirement might be for journals to require evidence of an a priori calculation via ethics application, trial registration or preregistration.}}

We completely agree. We have added the following text to capture this idea:

> Pages 17-18, Lines 281-285
>
> Even if power analysis reporting becomes the norm, it may remain unclear whether power analyses were conducted during the design of the study or at a later stage. Therefore, the most credible confirmatory studies will have preregistered power analyses. A bold journal might even require it.

\RC{\emph{Line 288: check reference format.}}

Thanks for catching this. We have fixed the issue.

<!-- # References -->

<!-- ::: {#refs custom-style="Bibliography"} -->
<!-- ::: -->

### 2022-10-19 docs/revision-1/r1_manuscript.Rmd
init

abstract: |
  *A priori* power analyses can ensure studies are unlikely to miss interesting effects. Recent metascience has suggested that kinesiology research may be underpowered and selectively reported. Here, we examined whether power analyses are being used to ensure informative studies in motor behavior. We reviewed every article published in three motor behavior journals between January 2019 and June 2021. Power analyses were reported in 13% of studies (*k* = 636) that tested a hypothesis. No study targeted the smallest effect size of interest. Most studies with a power analysis relied on estimates from previous experiments, pilot studies, or benchmarks to determine the effect size of interest. Studies without a power analysis reported support for their main hypothesis 85% of the time, while studies with a power analysis found support 76% of the time. The median sample sizes were *n* = 17.5 without a power analysis and *n* = 16 with a power analysis, suggesting the typical study design was underpowered for all but the largest plausible effect size. At present, power analyses are not being used to optimize the informativeness of motor behavior research. Adoption of this widely recommended practice may greatly enhance the credibility of the motor behavior literature.

@@

Standardized effect sizes also have conventional benchmarks that researchers may rely on when designing studies. Recent metascience suggests doing so is likely to result in underpowered research designs in practice [@Lovakov2021]. Instead of relying on benchmarks, some researchers may base their effect size target on a previous study or the results of a pilot study. However, large multi-lab replication studies have revealed that original studies may overestimate the true effect of an independent variable by 100% to 400% [@Klein2018; @OPENSCIENCECOLLABORATION2015]. Pilot studies are often even less helpful, as they tend to be smaller than published experiments so their estimates are even more imprecise [@Albers2018; @Kraemer2006; @Lakens2014]. When available, meta-analyses provide an effect size estimate based on the aggregation of available data. However, selective reporting of results can distort meta-analytic estimates and it can be difficult to correct for reporting bias [@Carter2019; @Thornton2000]. Nevertheless, estimates that have been corrected for reporting bias are more accurate than naïve random effects estimates and should be used when available [@Carter2019].

@@

Our design and analysis plan was preregistered after piloting our methods on a subsample of 40 papers. The preregistration, data, and code for this study is available using either of the following links: https://osf.io/wsdpv/ \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.

@@

Calculating *a priori* power for this study required estimating the final sample size and proportion of studies that would include a power analysis. Based on the pilot sample of articles, we estimated that the total number of studies we would extract would be approximately 500. The actual number was 636. We reasoned that if 10% of those studies included a power analysis, we would have 50 studies with power analyses and 450 studies without power analyses in our sample. The actual numbers were 13%, 85 and 551. Based on our rough estimates, we conducted simulations to estimate our power to detect differences in positive result rates of various plausible sizes. We based our expected positive results rate in studies without power analyses on estimates for psychology overall at 91.5% [@Fanelli2010]. We observed that, if our estimates were accurate, we would have 90% power to detect a difference of 16.5%, or a positive result rate of 75% in experiments with power analyses. Similarly, we estimated we would have 80% power to identify a positive result rate of 77.7% as significantly different. Unfortunately, if our estimated group sizes were accurate, we would have had low power (32%) to detect our smallest effect size of interest (6%). Given the actual sample sizes we observed, we had even greater power than planned to observe the effects we considered.

@@

All articles published in the *Journal of Motor Learning and Development*, *Human Movement Science*, and the *Journal of Motor Behavior* between January 2019 to June 2021 were uploaded to Covidence systematic review software and screened for inclusion (Figure \@ref(fig:fig1)). In total, 704 articles were reviewed. To be included in the analysis, studies were required to meet the following criteria: a) must be a primary study, b) must test a hypothesis, including the null hypothesis, c) there must be sufficient information available to adequately evaluate the criteria, and d) we must have access to the full-text. From the original 704 articles, 607 articles included at least one study that met the inclusion criteria and were included in the final analysis. Ninety-seven articles were excluded from the analysis for the following reasons: a) the studies were not primarily quantitative (63 studies), b) the studies made no hypothesis (27 studies), or there was insufficient information or a faulty DOI to assess the paper (7). The 607 included articles contributed a total of 636 eligible studies to the analysis.

@@

## Statistical Analysis

@@

We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target. A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results. A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results. Lastly, we conducted a two-tailed Welch’s *t*-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses. Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function [@R-rogme; @Rousselet2020; @Wilcox2021].

@@

## Proportion of Studies with a Power Analysis

Out of 636 total studies, 85 included a power analysis and 551 did not. Therefore, 13% of all studies sampled reported the results of a power analysis.

## Difference in Positivity Rates between Studies with and without a Power Analysis

As shown in Figure \@ref(fig:fig2), studies that did not include a power analysis reported finding support for their primary hypothesis 85% of the time (95% CI [82%, 88%]), while studies that included a power analysis found support 76% of the time (95% CI [66%, 85%]). The difference in positivity rates was not statistically significant, $\chi^2$= 3.47, *df* = 1, *p* = .06. The difference is positivity rates was not significantly smaller than our smallest effect size of interest, *Z* = .546, *p* = .71.

@@

## Justifications for Effect Sizes used in Power Analyses

The most common justification reported in our sample was to base the expected effect size on a previous study (*n* = 37), accounting for 44% of all justifications. The second most common justification was to provide no justification at all (*n* = 20), which occurred in 24% of studies that included a power analysis. Cohen’s benchmarks for small, medium, and large effects (*n* = 19) were used in 22% of studies. Pilot studies (*n* = 9) were used as justification in 11% of the sample.

## Power Levels Targeted in Power Analyses

The most frequently targeted power was 80%, which was chosen in 65% of studies with a power analysis (*n* = 55). The next most common power target was 95%, accounting for 14% of all power targets (*n* = 12); followed by 90% power, occurring in 11% of power analyses (*n* = 9). Two studies did not state their targeted power and several idiosyncratic power targets (96.7%, 96%, 95.33%, 85%, 75%, 70%, and 20%) were reported only once.

## Difference in Positivity Rates as a Function of Effect Size Justification

Figure \@ref(fig:fig3) illustrates the proportion of positive results for the four effect size justifications we found in our sample. Positivity rates were 100% for pilot study justification (9/9), 90% for studies with no justification (18/20), 68% for studies based on benchmarks (13/19), and 68% for studies based on previous studies (25/37). There was no significant difference between the positivity rates of any two justifications, $\chi^2$ = 7.12, *df* = 3, *p* = .068.

@@

## Difference in Positivity Rates as a Function of Target Power

Studies that targeted 80% power found support for their hypotheses 68% of the time (38/56). Studies that aimed for 90% power found support 100% of the time (8/8) and studies that aimed for 95% power found support 75% of the time (9/12). All studies that set an idiosyncratic power target or no target at all found support for their hypotheses (10/10). There was no significant difference between target power values, $\chi^2$= 7.22, *df* = 10, *p* = .70.

## Difference in Sample Size between Studies with and without a Power Analysis

Studies that included a power analysis had significantly smaller mean sample sizes (*M* = 21.91) than studies that did not include a power analysis (*M* = 40.98), *t*(624.59)= 3.43, *p* = .001. However, sample sizes were highly skewed---especially among studies without a power analysis. The median sample for those studies (*Mdn* = 17.5) was similar to the sample sizes of studies with a power analysis (*Mdn* = 16). We conducted a shift function as a sensitivity analysis and the results indicated no significant difference in sample size between studies with and without power analyses at any decile of their distributions.

@@

The purpose of this study was to investigate how frequently power analyses are reported in motor behavior articles, the justifications used for the effect size estimates, and the relationship between reporting power analyses and reporting positive results. We reviewed every article published in the *Journal of Motor Behavior*, the *Journal of Motor Learning and Development*, and *Human Movement Science* between January 2019 and June 2021 and identified 636 studies that tested a hypothesis. Of those 636 studies, 85 of them included a power analysis (13%). The rate of positive results was 85% overall and 76% when a power analysis was reported. The positive result rate was not significantly different between various effect size justifications or power targets.

Our results suggest that motor behavior research has not yet widely adopted power analyses to inform study design. When power analyses were reported, we observed a range of suboptimal effect size justifications. For example, 63% of studies that reported a power analysis based their effect size assumption on a previous study, a pilot study, or on effect size benchmarks. Another 24% of studies provided no justification at all. Each of these justifications (or lack thereof) are undesirable for different reasons. Previous studies---and especially pilot studies---are likely to provide exaggerated or noisy estimates of the unknown true effect. Effect size benchmarks may not match well the typical effect sizes one may find in their respective research area [@Lovakov2021]. Further, Cohen’s [@Cohen1988] benchmarks differ depending on which analysis is used in a power analysis. A medium effect is over twice as large for a multiple regression analysis as compared to a *t*-test [see @Correll2020 for a discussion with additional examples]. Not one study in the sample performed a power analysis based on their smallest effect size of interest. Power analyses can be an effective tool for researchers to ensure their studies are not underpowered, but to do so the smallest effects of interest need to be targeted.

The rate of positive results observed in this study suggest that positive findings are overrepresented in the motor behavior literature. While the studies in our sample reported positive results 84% of the time, the median per group sample size was ~17, which would provide 84% power to detect *d* = 1.05 with an independent *t*-test or *d* = .76 with a dependent *t*-test. In comparison, the most optimistic estimates for well-known motor behavior phenomena are much smaller. For example, the effect of feedback frequency on motor performance [*d* = .19, @McKay2022], self-controlled practice on retention performance [*d* = .54, @McKay2021], enhanced expectancies on retention performance [*d* = .54, @Bacelar2022], and external focus of attention on retention performance [*d* = .58, @Chua2021]. Estimates for the true effects of these phenomena that have been corrected for reporting-bias are markedly smaller, ranging from *d* = 0 to *d* = .25. Assuming the average effects investigated by the studies in our sample were similar to the optimistic estimates for other motor behavior effects, it is likely this literature was underpowered on average and potentially heavily censored.

All three journals that we sampled from either explicitly mention power in their instructions for authors (*Human Movement Science*), or reference JARS (*Journal of Motor Learning and Development*) or CONSORT (*Journal of Motor Behavior*) reporting standards; both of which include power analyses. Therefore, adoption of power analyses targeting interesting effects does not require a policy shift, simply the enforcement of current guidelines. Since no studies in this sample powered for the smallest effect size of interest, if *Human Movement Science*, the *Journal of Motor Behavior*, and/or the *Journal of Motor Learning and Development* enforce their existing guidelines then their future publications will look dramatically different. We believe this is a promising path forward to increase the reliability of motor behavior research and the evidence-based recommendations for coaching and rehabilitation.

@@

@@

<!-- ## Author Contributions (CRediT Taxonomy) -->

>>>>

abstract: |
  *A priori* power analyses can ensure studies are unlikely to miss interesting effects. Recent metascience has suggested that kinesiology research may be underpowered and selectively reported. Here, we examined whether power analyses are being used to ensure informative studies in motor behavior. We reviewed every article published in three motor behavior journals between January 2019 and June 2021. Power analyses were reported in 13% of studies (*k* = \textcolor{blue}{635}) that tested a hypothesis. No study targeted the smallest effect size of interest. Most studies with a power analysis relied on estimates from previous experiments, pilot studies, or benchmarks to determine the effect size of interest. Studies without a power analysis reported support for their main hypothesis 85% of the time, while studies with a power analysis found support \textcolor{blue}{77}% of the time. The median sample sizes were *n* = 17.5 without a power analysis and *n* = 16 with a power analysis, suggesting the typical study design was underpowered for all but the largest plausible effect size. At present, power analyses are not being used to optimize the informativeness of motor behavior research. Adoption of this widely recommended practice may greatly enhance the credibility of the motor behavior literature.

@@

Standardized effect sizes also have conventional benchmarks that researchers may rely on when designing studies. Recent metascience suggests doing so is likely to result in underpowered research designs in practice [@Lovakov2021]. Instead of relying on benchmarks, some researchers may base their effect size target on a previous study or the results of a pilot study. However, large multi-lab replication studies have revealed that original studies may overestimate the true effect of an independent variable by 100% to 400% [@Klein2018; @OPENSCIENCECOLLABORATION2015]. Pilot studies are often even less helpful, as they tend to be smaller than published experiments so their estimates are even more imprecise [@Albers2018; @Kraemer2006; @Lakens2014]. When available, meta-analyses provide an effect size estimate based on the aggregation of available data. However, selective reporting of results can distort meta-analytic estimates and it can be difficult to correct for reporting bias [@Carter2019; @Thornton2000]. Nevertheless, estimates that have been corrected for reporting bias are more accurate than naïve random effects estimates and should be used when available [@Carter2019]. \textcolor{blue}{Targeting the estimated mean effect will only provide the desired power if the effect is average or above average in the specific context being investigated. Therefore, even if researchers are privy to a precise and unbiased estimate of the mean effect they are studying, they still might choose a different effect to target in their power analysis.}

@@

Our design and analysis plan was preregistered after piloting our methods on a subsample of 40 papers. The preregistration, materials, data, and code are available using either of the following links: https://osf.io/wsdpv/ \censor{or} \censor{$https://github.com/cartermaclab/proj_power-motor-behaviour$}.

@@

Calculating *a priori* power for this study required estimating the final sample size and proportion of studies that would include a power analysis. Based on the pilot sample of articles, we estimated that the total number of studies we would extract would be approximately 500. The actual number was \textcolor{blue}{635}. We reasoned that if 10% of those studies included a power analysis, we would have 50 studies with power analyses and 450 studies without power analyses in our sample. The actual numbers were 13%, \textcolor{blue}{84} and 551. Based on our rough estimates, we conducted simulations to estimate our power to detect differences in positive result rates of various plausible sizes. We based our expected positive results rate in studies without power analyses on estimates for psychology overall at 91.5% [@Fanelli2010]. We observed that, if our estimates were accurate, we would have 90% power to detect a difference of 16.5%, or a positive result rate of 75% in experiments with power analyses. Similarly, we estimated we would have 80% power to identify a positive result rate of 77.7% as significantly different. Unfortunately, if our estimated group sizes were accurate, we would have had low power (32%) to detect our smallest effect size of interest (6%). Given the actual sample sizes we observed, we had even greater power than planned to observe the effects we considered.

@@

All articles published in the *Journal of Motor Learning and Development*, *Human Movement Science*, and the *Journal of Motor Behavior* between January 2019 to June 2021 were uploaded to Covidence systematic review software and screened for inclusion (Figure \@ref(fig:fig1)). In total, 704 articles were reviewed. To be included in the analysis, studies were required to meet the following criteria: a) must be a primary study, b) must test a hypothesis, including the null hypothesis, c) there must be sufficient information available to adequately evaluate the criteria, and d) we must have access to the full-text. From the original 704 articles, \textcolor{blue}{606} articles included at least one study that met the inclusion criteria and were included in the final analysis. Ninety-\textcolor{blue}{eight} articles were excluded from the analysis for the following reasons: a) the studies were not primarily quantitative (63 studies), b) the studies made no hypothesis (27 studies), or there was insufficient information or a faulty DOI to assess the paper (\textcolor{blue}{8}). The \textcolor{blue}{606} included articles contributed a total of \textcolor{blue}{635} eligible studies to the analysis.

@@

## Statistical analysis

### Pre-registered analyses

@@

We calculated the percentage of studies that conducted a power analysis with a) each effect justification and b) each power target. A two-sided, six sample proportion test was conducted to test whether at least two different effect size justifications in power analyses led to different rates of positive results. A two-sided, 11-sample proportion test was conducted to test whether at least two power targets resulted in a different rate of positive results. We conducted a two-tailed Welch’s *t*-test to determine whether studies with power analyses had different sample sizes compared to studies without power analyses. Given the data were highly skewed, we also conducted a sensitivity analysis using a shift function [@R-rogme; @Rousselet2020; @Wilcox2021].

### Exploratory analysis

\textcolor{blue}{We used a 3-sample proportion test to compare the proportion of studies including a power analysis in the three journals surveyed.}

@@

## Pre-registered analyses

### Proportion of studies with a power analysis

Out of \textcolor{blue}{635} total studies, \textcolor{blue}{84} included a power analysis and 551 did not\textcolor{blue}{; thus,} 13% of all studies sampled reported the results of a power analysis.

### Difference in positivity rates between studies with and without a power analysis

As shown in Figure \@ref(fig:fig2), studies that did not include a power analysis reported finding support for their primary hypothesis 85% of the time (95% CI [82%, 88%]), while studies that included a power analysis found support \textcolor{blue}{77}% of the time (95% CI [\textcolor{blue}{67}%, \textcolor{blue}{86}%]). The difference in positivity rates was not statistically significant, $\chi^2$= \textcolor{blue}{2.71}, *df* = 1, *p* = \textcolor{blue}{.10}. The difference is positivity rates was not significantly smaller than our smallest effect size of interest, *Z* = \textcolor{blue}{.361}, *p* = \textcolor{blue}{.641}.

@@

### Justifications for effect sizes used in power analyses

The most common justification reported in our sample was to base the expected effect size on a previous study (*n* = 37), accounting for 44% of all justifications. The second most common justification was to provide no justification at all (*n* = 20), which occurred in 24% of studies that included a power analysis. Cohen’s benchmarks for small, medium, and large effects (*n* = \textcolor{blue}{18}) were used in \textcolor{blue}{21}% of studies. Pilot studies (*n* = 9) were used as justification in 11% of the sample.

### Power levels targeted in power analyses

The most frequently targeted power was 80%, which was chosen in \textcolor{blue}{64}% of studies with a power analysis (*n* = \textcolor{blue}{54}). The next most common power target was 95%, accounting for 14% of all power targets (*n* = 12); followed by 90% power, occurring in 11% of power analyses (*n* = 9). Two studies did not state their targeted power and several idiosyncratic power targets (96.7%, 96%, 95.33%, 85%, 75%, 70%, and 20%) were reported only once.

### Difference in positivity rates as a function of effect size justification

Figure \@ref(fig:fig3) illustrates the proportion of positive results for the four effect size justifications we found in our sample. Positivity rates were 100% for pilot study justification (9/9), 90% for studies with no justification (18/20), \textcolor{blue}{72}% for studies based on benchmarks (13/\textcolor{blue}{18}), and 68% for studies based on previous studies (25/37). There was no significant difference between the positivity rates of any two justifications, $\chi^2$ = \textcolor{blue}{6.76}, *df* = 3, *p* = \textcolor{blue}{.08}.

@@

### Difference in positivity rates as a function of target power

Studies that targeted 80% power found support for their hypotheses \textcolor{blue}{70}% of the time (38/\textcolor{blue}{54}). Studies that aimed for 90% power found support 100% of the time (8/8) and studies that aimed for 95% power found support 75% of the time (9/12). All studies that set an idiosyncratic power target or no target at all found support for their hypotheses (10/10). There was no significant difference between target power values, $\chi^2$= \textcolor{blue}{6.82}, *df* = 10, *p* = \textcolor{blue}{.743}.

### Difference in sample size between studies with and without a power analysis

Studies that included a power analysis had significantly smaller mean sample sizes (*M* = \textcolor{blue}{21.75}) than studies that did not include a power analysis (*M* = 40.98), *t*(\textcolor{blue}{624.63})= \textcolor{blue}{3.45}, *p* = .001. However, sample sizes were highly skewed---especially among studies without a power analysis. The median sample for those studies (*Mdn* = 17.5) was similar to the sample sizes of studies with a power analysis (*Mdn* = 16). We conducted a shift function as a sensitivity analysis and the results indicated no significant difference in sample size between studies with and without power analyses at any decile of their distributions.

## Exploratory analysis

### Difference among journals in proportion of studies with a power analysis

\textcolor{blue}{Studies published in Human Movement Science included a power analysis (59/271) 18\% of the time, the Journal of Motor Learning and Development (9/128) 10\% of the time, and the Journal of Motor Behavior (16/152) 7\% of the time. The 3-samples proportions test was significant, $\chi^2$ = 13.52, \emph{df} = 2, \emph{p} = .001.}

@@

The purpose of this study was to investigate how frequently power analyses are reported in motor behavior articles, the justifications used for the effect size estimates, and the relationship between reporting power analyses and reporting positive results. We reviewed every article published in the *Journal of Motor Behavior*, the *Journal of Motor Learning and Development*, and *Human Movement Science* between January 2019 and June 2021 and identified 63\textcolor{blue}{5} studies that tested a hypothesis. Of those 63\textcolor{blue}{5} studies, 8\textcolor{blue}{4}  included a power analysis (13%). The rate of positive results was 85% overall and 7\textcolor{blue}{7}% when a power analysis was reported. The positive result rate was not significantly different between various effect size justifications or power targets.

Our results \textcolor{blue}{cause us to wonder whether} motor behavior research has not yet widely adopted power analyses to inform study design. When power analyses were reported, we observed a range of suboptimal effect size justifications. \textcolor{blue}{Specifically, 76\%} of studies that reported a power analysis based their effect size assumption on a previous study \textcolor{blue}{(\emph{n} = 37, 44\%)}, a pilot study  \textcolor{blue}{(\emph{n} = 9, 11\%)}, or on effect size benchmarks \textcolor{blue}{(\emph{n} = 18, 21\%)}. \textcolor{blue}{The remaining} 24% of studies provided no justification at all. Each of these justifications (or lack thereof) are undesirable for different reasons. Previous studies---and especially pilot studies---are likely to provide exaggerated or noisy estimates of the unknown true effect. Effect size benchmarks may not match well the typical effect sizes one may find in their respective research area [@Lovakov2021]. Further, Cohen’s [@Cohen1988] benchmarks differ depending on which analysis is used in a power analysis. A medium effect is over twice as large for a multiple regression analysis as compared to a *t*-test [see @Correll2020 for a discussion with additional examples]. Not one study in the sample performed a power analysis based on their smallest effect size of interest. Power analyses can be an effective tool for researchers to ensure their studies are not underpowered, but to do so the smallest effects of interest need to be targeted.

The rate of positive results observed in this study suggest that positive findings are overrepresented in the motor behavior literature. While the studies in our sample reported positive results 84% of the time, the median per group sample size was ~17, which would provide 84% power to detect *d* = 1.05 with an independent *t*-test or *d* = .76 with a dependent *t*-test. In comparison, the most optimistic estimates for well-known motor behavior phenomena are much smaller. For example, the effect of feedback frequency on motor performance [*d* = .19, @McKay2022], self-controlled practice on retention performance [*d* = .54, @McKay2021], enhanced expectancies on retention performance [*d* = .54, @Bacelar2022], and external focus of attention on retention performance [*d* = .58, @Chua2021]. Estimates for the true effects of these phenomena that have been corrected for reporting-bias are markedly smaller, ranging from *d* = 0 to *d* = .25. Assuming the average effects investigated by the studies in our sample were similar to the optimistic estimates for other motor behavior effects, it is likely this literature was underpowered on average and potentially heavily censored.[^1]

[^1]: \textcolor{blue}{We use the term "censored" here to mean that some experiments fail to enter the published literature for results-dependent reasons, commonly called publication bias or reporting bias.}

All three journals that we sampled from either explicitly mention power in their instructions for authors (*Human Movement Science*), or reference JARS (*Journal of Motor Learning and Development*) or CONSORT (*Journal of Motor Behavior*) reporting standards; both of which include power analyses. \textcolor{blue}{Perhaps the approach of more directly instructing authors to include a power analysis explains why Human Movement Science had a higher proportion of studies with a power analysis. This post-hoc explanation is consistent with the modest increase in power analysis reporting observed following adoption of guidelines requiring sample size justification in the journal Gait \& Posture} [@McCrum2022]. \textcolor{blue}{ However, less than one in five studies in \emph{Human Movement Science} included a power analysis so including explicit instructions to authors is not sufficient to promote widespread reporting. It seems the} adoption of power analyses targeting interesting effects does not require a policy shift, simply the enforcement of current guidelines. Since no studies in this sample powered for the smallest effect size of interest, if *Human Movement Science*, the *Journal of Motor Behavior*, and/or the *Journal of Motor Learning and Development* enforce their existing guidelines then their future publications \textcolor{blue}{might} look dramatically different. \textcolor{blue}{Even if power analysis reporting becomes the norm, it may remain unclear whether power analyses were conducted during the design of the study or at a later stage. Therefore, the most credible confirmatory studies will have preregistered power analyses. A bold journal might even require it.} We believe this is a promising path forward to increase the reliability of motor behavior research and the evidence-based recommendations for coaching and rehabilitation.

@@

\textcolor{blue}{Selection bias makes it challenging to interpret the impact of power analyses on positive results and sample sizes. It is possible that power analyses are sometimes added to a study to defend a null result or a small sample size. It is also possible that researchers who conduct power analyses choose to study larger effects, leading to smaller sample sizes. We encourage readers to focus on the overall low frequency of power analysis reporting, small average sample sizes, and high rates of positive results as a description of the modern motor behavior literature.}

@@

## Author Contributions (CRediT Taxonomy)
